Efficient Handling of String-Number Conversion

### 2019-11-21 draft.tex
aaaarghh

For many languages such as Java, string-number conversions are core operations, and much symbolic execution can not been done without supporting it. In fact,  a code that receives input tends to need to convert at least some of that input into numbers,  which makes the handling of string-integer conversions of the utmost importance. For example, the program fragment below is a variant of the Luhn test algorithm that is often used in credit card or personal ID validation.

@@

The input \verb|value| of the Luhn test algorithm is a sequence of digits. The algorithm processes the digits in the reversed order. The value of every odd digit (e.g., 1st, 3rd, etc.) is added to \verb|sum| directly. For the value of every even digit, the algorithm (1) doubles its value, (2) subtracts its value by $9$ if the doubled-value is larger than $9$, and (3) adds the final result to \verb|sum|. At the end, the input is validated if the string value of \verb|sum| ends by an $0$ (i.e., \verb| sum mod 10=0|).

To check whether the program path that traverses both loops exactly once, which corresponds to the case when $| \verb|value||=2$, and finally passes this test has a valid input, we create the following (string) constraint:

@@

The semantics of a string constraint $\phi$ restricted by $\dr$ is then defined as 
$\semof{\phi}^\dr = \{I\in\semof \phi \mid \forall x\in X:I(x)\in\semof{\dr(x)}\} \subseteq \vars\times \varn \rightarrow \Sigma^* \times\nat$.
The following sections will be devoted to describing the flattening of the atomic string constraints. 
The correctness of the entire construction of $\outlf$ is expressed by  Theorem~\ref{theorem:correct}:
\begin{theorem}\label{theorem:correct}
$\decode R (\semof {\outlf}) = \semof{\insc}^\dr$ 
\end{theorem}

@@

\changed{
To formalize this correctness argument, we need the notion of 
%
an \emph{$\dr$-encoding} of an interpretation $I \cup I'$ of a general string constraint, consisting of an interpretation $I$ of string variables and an interpretation $I'$ of integer variables:
%
An $\dr$-encoding of $I \cup I'$ is a union $e \cup I'$ of the original integer interpretation $I'$ and an interpretation $e : V_\dr \cup \#V_\dr \rightarrow \nat$ that encodes $I$, that is, it satisfies that for each string variable $x$ of $I$, $\decode {\dr(x)} (e_{V_x\cup\#V_x}) = I(x)$.  
}

The set of all $\dr$-encodings of $(I\cup I')$ is then denoted $\encode \dr {(I\cup I')}$
%\yfc{need a serious rewrite :-) Maybe I need to know why you define so many encodings...}
%
%Further, if $I \cup I'$ is an interpretation of a general string constraint, that is, an interpretation $I$ of string variables $X$ united with an interpretation $I'$ of integer variables $N$,
%%
%then an $R$-encoding of $I \cup I'$ is any interpretation $e \cup I$ with  $e:V_R \cup \#V_R \cup \Simga_\epsilon$ 
%such that $\encode R {(I\cup I')} \defeq \{e \cup I' \mid e\in \encode R(I)\}$ of interpretations of variables $V_R \cup \#V_R \cup N$. That is, the interpretation of string variables is encoded and the interpretation of integer variables is kept as is.
%
%
and we lift this notation to sets of interpretation in the standard manner, i.e., $\encode \dr (J) = \bigcup_{I \in J}\encode \dr (I)$ for a set of interpretations $J$.
%
The inductive argument needed in the correctness proof of the under-approximation then reads as
$$\semof{\underf \dr \phi}_{V_\dr\cup V_{\#\dr}}  = \encode \dr{\semof{\phi}}$$ 
%
In the next sections, 
we will formulate the corresponding correctness lemma for flattening constructed from each type of atomic string constraints.
%
We note that the restriction of the semantics of $\underf \dr \phi$ here is needed since $\underf \dr \phi$ will be constructed with some auxiliary variables.

>>>>

For many languages such as Java, string-number conversions are core operations, and much symbolic execution can not been done without supporting it. In fact,  a code that receives input tends to need to convert at least some of that input into numbers,  which makes the handling of string-integer conversions of the utmost importance. For example, the program fragment below is a variant of the Luhn test algorithm \cite{?} that is often used in credit card or personal ID validation.

@@

The input \verb|value| of the Luhn test algorithm is a sequence of digits. The algorithm process the digits in the reversed order. The value of every odd digit (e.g., 1st, 3rd, etc.) is added to \verb|sum| directly. For the value of every even digit, the algorithm (1) doubles its value, (2) subtracts its value by $9$ if the doubled-value is bigger than $9$, and (3) adds the final result to \verb|sum|. At the end, the input is validated if the string value of \verb|sum| ends by an $0$ (i.e., \verb| sum mod 10=0|).

To check whether the program path that traverses both loop exactly once, which corresponds to the case when $| \verb|value||=2$, and finally passes this test has a valid input, we create the following (string) constraint:

@@

The semantics of a string constraint $\phi$ restricted by $\dr$ is then defined as 
$\semof{\phi}^\dr = \{I\in\semof \phi \mid \forall x\in X:I(x)\in\semof{\dr(x)}\} \subseteq \vars\times \varn \rightarrow \Sigma^* \times\nat$.
The following sections will be devoted to describing the flattening of the atomic string constraints.

The correctness of the entire construction of $\outlf$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterised by the domain restriction $\dr$. 
For an interpretation $e \cup I'$ consisting of an interpretaton $e$ of $V_\dr$ and $\#V_\dr$ and an interpretation fo integer variavles $I'$, we let $\decode \dr (e\cup I') = (I \cup I')$ where $I$ is the interpretation of string variables such that for each $x\in X$,
$I(x) = \decode {\dr(x)} (e_{V_{\dr(x)}\cup \#V_{\dr(x)}})$. That is, $\decode \dr (e\cup I')$ returns the original integer interpretation $I'$ together with each string variable $x$ interpreted as the $\decode {\dr(x)}$-decoding of the values of $V_{\dr(x)}$ and $\#V_{\dr(x)}$. 
We lift $\decode \dr$ to sets of interpretations in the standard manner.
\begin{theorem}\label{theorem:correct}
$\decode R (\semof {\outlf}) = \semof{\insc}^\dr$ 
\end{theorem}

@@

%To formalize this correctness argument, we need the notion of 
%%
%an \emph{$\dr$-encoding} of an interpretation $I \cup I'$ of a general string constraint, consisting of an interpretation $I$ of string variables and an interpretation $I'$ of integer variables:
%%
%An $\dr$-encoding of $I \cup I'$ is a union $e \cup I'$ of the original integer interpretation $I'$ and an interpretation $e : V_\dr \cup \#V_\dr \rightarrow \nat$ that encodes $I$, that is, it satisfies that for each string variable $x$ of $I$, $\decode {\dr(x)} (e_{V_x\cup\#V_x}) = I(x)$.

To formalize this correctness argument, we define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a set $J$ of string/integer variables interpretations,  
we let $\encode \dr (J) \defeq \{e\mid \decode \dr (e) \in J\}$.
%\yfc{need a serious rewrite :-) Maybe I need to know why you define so many encodings...}
%
%Further, if $I \cup I'$ is an interpretation of a general string constraint, that is, an interpretation $I$ of string variables $X$ united with an interpretation $I'$ of integer variables $N$,
%%
%then an $R$-encoding of $I \cup I'$ is any interpretation $e \cup I$ with  $e:V_R \cup \#V_R \cup \Simga_\epsilon$ 
%such that $\encode R {(I\cup I')} \defeq \{e \cup I' \mid e\in \encode R(I)\}$ of interpretations of variables $V_R \cup \#V_R \cup N$. That is, the interpretation of string variables is encoded and the interpretation of integer variables is kept as is.
%
%
%and we lift this notation to sets of interpretation in the standard manner, i.e., $\encode \dr (J) = \bigcup_{I \in J}\encode \dr (I)$ for a set of interpretations $J$.
%
The inductive argument needed in the correctness proof of the under-approximation then reads as
$$\semof{\underf \dr \phi}_{V_\dr\cup V_{\#\dr}}  = \encode \dr{\semof{\phi}}$$ 
%
In the next sections, 
we will formulate the corresponding correctness lemma for flattening constructed from each type of atomic string constraints.
%
We note that the restriction of the semantics of $\underf \dr \phi$ here is needed since $\underf \dr \phi$ will be constructed with some auxiliary variables.
